---
title: "GWAS report"
author: "Rodrigo Bonazzola"
output:
  html_document:
    theme: flatly
    code_folding: hide
runtime: shiny
---

```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(glue)
```
This report aims to summarise the findings after following Bernard's suggestions.

### Paths
```{r}
data_dir <- "~/data/PhD/UKBB/"
GWAS_DIR <- "/home/rodrigo/GWAS/output/coma"

gwas_paths <- list(
  "qqplot" = file.path(GWAS_DIR, "{input$experiment}/{input$suffix}/figures/GWAS__{as.character(input$z)}__{input$suffix}__QQ-plot.png"),
  "manhattan" = file.path(GWAS_DIR, "{input$experiment}/{input$suffix}/figures/GWAS__{as.character(input$z)}__{input$suffix}__manhattan.png"),
  "pooled_qqplot" = file.path(GWAS_DIR, "{input$experiment}/{input$suffix}/figures/GWAS__all__QQ-plot.png")
)

```

## Filtering GWAS sample
### Loading subject IDs
#### British individuals
Only British individuals were kept, based on self-reported ethnicity (data field `21000`, value `1001`).
```{r}
gbr_ids_file <- glue("{data_dir}/subject_ids/british_ids.txt")
cmr_gbr_ids_file <- glue("{data_dir}/subject_ids/cmr_british_ids.txt")
```

```{r}
gbr_ids <- read.csv(gbr_ids_file, header = FALSE)$ID
ids <- read.csv(glue("{data_dir}/subject_ids/IDs_32k.csv"), header = FALSE)$V1
```

```{r}
cardiac_indices_df <-  read.csv(glue("{data_dir}/cardiac_indices/Cardiac_Function_Indexes_11350.csv"))
```

#### EF < 50%
```{r}
read_icd10_ids <- function(icd10) {
  read.csv(glue("{data_dir}/subject_ids/icd10/{icd10}.txt"))[,1]
}
```

#### Hypertension
```{r}
hypertension_ids_all <- read_icd10_ids("I10")
hypertension_ids_cmr <- intersect(ids, hypertension_ids_all)
```

#### Acute myorcardial infarction
```{r}
icd10_codes_acuteMI <- c(
  "I210", # 
  "I211", #
  "I212", #
  "I213", #
  "I214", #
  "I219"  #
)
acuteMI_ids_all <- unlist(sapply(icd10_codes_acuteMI, read_icd10_ids))
acuteMI_cmr <- intersect(ids, acuteMI_ids_all)
```

#### Cardiomyopathies
```{r}
icd10_codes_CM <- c(
  "I420", #
  "I421", #
  "I422", #
  "I424", #
  "I425", #
  "I426", #
  "I427", #
  "I428", #
  "I429"  #
)

CM_ids_all <- unlist(sapply(icd10_codes_CM, read_icd10_ids))
CM_cmr <- intersect(ids, CM_ids_all)
```

#### Angina Pectoris
```{r}
icd10_codes_AP <- c(
  "I200", #
  "I201", #
  "I208", #
  "I209"  #
)

AP_ids_all <- unlist(sapply(icd10_codes_AP, read_icd10_ids))
AP_cmr <- intersect(ids, AP_ids_all)
```

`r length(hypertension_ids_cmr)` British individuals also with CMR.

`r length(hypertension_ids_all)` individuals with hypertension.
Among them, `r length(hypertension_ids_cmr)` British individuals also with CMR.

`r length(acuteMI_ids_all)` individuals with acute MI.
Among them, `r length(acuteMI_cmr)` British individuals also with CMR.

`r length(CM_ids_all)` individuals with cardiomyopathies.
Among them, `r length(CM_cmr)` British individuals also with CMR.

`r length(AP_ids_all)` individuals with angina pectoris.
Among them, `r length(AP_cmr)` British individuals also with CMR.

#### Relatedness
Kinship coefficients were obtained from the UK Biobank, by running the command `ukbgene rel`.

```{r}
kinship_df <- read.table(glue("{data_dir}/ukb11350_rel_s488226.dat"), header = TRUE)
kinship_df <- kinship_df %>% filter(ID1 %in% ids & ID2 %in% ids)
str(kinship_df)
```


### GWAS results
```{r}
experiments <- c(
  "scaled"="2020-09-11_02-13-41",  # scaled meshes
  "unscaled"="2020-09-30_12-36-48"   # unscaled meshes
)

run_params <- yaml::read_yaml("/home/rodrigo/pytorch_coma/analysis/paper_z_mapping_parameters.yaml")
# mapping <- run_params[[run_id]]$mapping
# sign <- run_params[[run_id]]$sign
suffixes <- list.files(file.path(GWAS_DIR, experiments["scaled"])) %>% .[grepl("std", .)]
zs <- paste0("z", 0:7)

filtering <- c(
  "No filtering"="std_covariates__GBR__qc",
  "EF < 50%"="std_covariates__GBR__excl_EF_lt_50__qc",
  "EF < 50% + HT"="std_covariates__GBR__excl_EF_lt_50_and_hypertension__qc",
  "EF < 50% + HT + CM"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI__qc",
  "EF < 50% + HT + CM + acMI"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI_Cardiomyopathies__qc",
  "EF < 50% + HT + CM + acMI + AP"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI_Cardiomyopathies_AnginaPectoris__qc"
)
```

```{r, echo=FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
        selectInput("experiment", "Experiment:", choices = experiments, selected = "2020-09-11_02-13-41"),
        uiOutput("zSelection"),
        selectInput("suffix", "Filtering:", choices = filtering),
        radioButtons("plotType", "Plot:", choices = c("Manhattan"="manhattan", "QQ plot"="qqplot"))
      ),
      mainPanel(
        conditionalPanel("input.plotType == 'manhattan'",
          imageOutput("manhattan")
        ),
        conditionalPanel("input.plotType == 'qqplot'",
          imageOutput("qqplot")
        )
      )
    )
  ),

  server = function(input, output) {
    output$zSelection <- renderUI({
      run_params <- yaml::read_yaml("/home/rodrigo/pytorch_coma/analysis/paper_z_mapping_parameters.yaml")
      # mapping <- run_params[[if_else(is.null(input$experiment), "scaled", input$experiment)]]$mapping
      print(input$experiment)
      mapping <- run_params[[input$experiment]]$mapping
      zs_ <- paste0("z", 1:8)
      names(zs_) <- zs_
      zs_ <- sapply(zs_ , function(x) mapping[[x]])
      print(zs_)
      selectInput("z",  "z:", choices = zs_)
    })
    
    output$manhattan <- renderImage( { 
      list(
        src = glue(gwas_paths$manhattan),
        width = 600, height = 200
      ) 
    }, deleteFile=FALSE )
    
    output$qqplot <- renderImage( { 
      list(
        src = glue(gwas_paths$qqplot),
        width = 300, height = 300
      ) 
    }, deleteFile=FALSE )
  }
)
```