---
title: "GWAS report"
author: "Rodrigo Bonazzola"
output:
  html_document:
    theme: flatly
    code_folding: hide
---
  
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(glue)

data_dir <- "~/data/PhD/UKBB/"
GWAS_DIR <- "/home/rodrigo/GWAS/output/coma"

gwas_paths <- list(
  "qqplot" = file.path(GWAS_DIR, "{experiment}/{suffix}/figures/GWAS__{as.character(z)}__{suffix}__QQ-plot.png"),
  "manhattan" = file.path(GWAS_DIR, "{experiment}/{suffix}/figures/GWAS__{as.character(z)}__{suffix}__manhattan.png") #,
  # "pooled_qqplot" = file.path(GWAS_DIR, "{input$experiment}/{input$suffix}/figures/GWAS__all__QQ-plot.png")
)
```

This report aims to summarise the findings after following Bernard's suggestions.

## Filtering GWAS sample

Only British individuals were kept, based on self-reported ethnicity (data field `21000`, value `1001`).
In order to identify related individuals, kinship coefficients were obtained from the UK Biobank, by running the command `ukbgene rel`.
```{r}
gbr_ids_file <- glue("{data_dir}/subject_ids/british_ids.txt")
cmr_gbr_ids_file <- glue("{data_dir}/subject_ids/cmr_british_ids.txt")
gbr_ids <- read.csv(gbr_ids_file, header = TRUE)$ID
ids <- read.csv(glue("{data_dir}/subject_ids/IDs_32k.csv"), header = FALSE)$V1

kinship_df <- read.table(glue("{data_dir}/ukb11350_rel_s488226.dat"), header = TRUE)
kinship_df <- kinship_df %>% filter(ID1 %in% ids & ID2 %in% ids)
ids_related <- as.character(kinship_df$ID1)
```

There were `r length(gbr_ids)` British individuals in the full cohort, and from them `r length(setdiff(intersect(ids, gbr_ids), ids_related))` unrelated individuals with CMR were processed.

### LVEF < 50%
```{r}
cardiac_indices_df <-  read.csv(glue("{data_dir}/cardiac_indices/Cardiac_Function_Indexes_11350.csv"))
hist(cardiac_indices_df$LVEF_automatic, main="Distribution of LV ejection fraction", xlab="LVEF")

n_ef_lt_50 <- length(cardiac_indices_df$LVEF_automatic[cardiac_indices_df$LVEF_automatic < 50])
```

`r n_ef_lt_50` subjects with LVEF < 50%.


### Diagnoses
```{r, echo=FALSE}
read_icd10_ids <- function(icd10) {
  read.csv(glue("{data_dir}/subject_ids/icd10/{icd10}.txt"))[,1]
}
```


```{r}
hypertension_ids_all <- read_icd10_ids("I10")
hypertension_ids_cmr <- intersect(ids, hypertension_ids_all)
```
#### Hypertension (HT)
- `r length(hypertension_ids_all)` individuals with hypertension.
Among them, `r length(hypertension_ids_cmr)` British individuals also with CMR.

```{r}
icd10_codes_acuteMI <- c(
  "I210", # 
  "I211", #
  "I212", #
  "I213", #
  "I214", #
  "I219"  #
)
acuteMI_ids_all <- unlist(sapply(icd10_codes_acuteMI, read_icd10_ids))
acuteMI_cmr <- intersect(ids, acuteMI_ids_all)
```
#### Acute myorcardial infarction (AcMI)
- `r length(acuteMI_ids_all)` individuals with acute MI.
Among them, `r length(acuteMI_cmr)` British individuals also with CMR.

```{r}
icd10_codes_CM <- c(
  "I420", #
  "I421", #
  "I422", #
  "I424", #
  "I425", #
  "I426", #
  "I427", #
  "I428", #
  "I429"  #
)

CM_ids_all <- unlist(sapply(icd10_codes_CM, read_icd10_ids))
CM_cmr <- intersect(ids, CM_ids_all)
```
#### Cardiomyopathies (CM)
- `r length(CM_ids_all)` individuals with cardiomyopathies.
Among them, `r length(CM_cmr)` British individuals also with CMR.

```{r}
icd10_codes_AP <- c(
  "I200", #
  "I201", #
  "I208", #
  "I209"  #
)

AP_ids_all <- unlist(sapply(icd10_codes_AP, read_icd10_ids))
AP_cmr <- intersect(ids, AP_ids_all)
```
#### Angina Pectoris (AP)
- `r length(AP_ids_all)` individuals with angina pectoris.
Among them, `r length(AP_cmr)` British individuals also with CMR.


### GWAS results


```{r myfile-1-plot}

filtering <- c(
  "No filtering"="std_covariates__GBR__qc",
  "EF < 50%"="std_covariates__GBR__excl_EF_lt_50__qc",
  "EF < 50% + HT"="std_covariates__GBR__excl_EF_lt_50_and_hypertension__qc",
  "EF < 50% + HT + CM"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI__qc",
  "EF < 50% + HT + CM + acMI"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI_Cardiomyopathies__qc",
  "EF < 50% + HT + CM + acMI + AP"="std_covariates__GBR__excl_EF_lt_50_and_hypertension_acuteMI_Cardiomyopathies_AnginaPectoris__qc"
)

```

### Unscaled meshes: $z_{\text{LVV}}$
```{r, echo = F, results = 'asis'}
experiment <- "2020-09-30_12-36-48"
for (i in 1:length(filtering)) {
  z = "z0"
  caption <- names(filtering)[i]
  suffix <- filtering[i]
  cat(paste0('\n#### ', caption, ":"))
  cat(paste0('\n![', '](', glue(gwas_paths$manhattan), ')\n\n'))
}
```

### Scaled meshes: $z_{\text{LVSph}}$
```{r, echo = F, results = 'asis'}
experiment <- "2020-09-11_02-13-41"

for (i in 1:length(filtering)) {
  z = "z5"
  caption <- names(filtering)[i]
  suffix <- filtering[i]
  cat(paste0('\n#### ', caption))
  cat(paste0('\n![', '](', glue(gwas_paths$manhattan), ')\n\n'))
}
```

